---
title: "covid19-philadelphia"
output: github_document
---

### About this repo
De-identified, aggregate datasets showing COVID-19 cases, hospitalizations and deaths by date, zip, or age/sex/race as made available by the City of Philadelphia through its Open Data Program.

The data in this repository is a history of daily snapshots made available here: https://www.opendataphilly.org/showcase/philadelphia-covid-19-information

The following datasets have been collected since 6/4/2020:

* COVID Cases by Age
* COVID Cases by Date
* COVID Cases by Sex
* COVID Cases by Zipcode
* COVID Deaths by Age
* COVID Deaths by Date
* COVID Deaths by Sex

These have been collected since 8/14/2020:

* COVID Cases by Race
* COVID Deaths by Race
* COVID Hospitalizations by Age
* COVID Hospitalizations by Date
* COVID Hospitalizations by Race
* COVID Hospitalizations by Sex
* COVID Hospitalizations by Zipcode

Starting on 9/29/2020, COVID Cases by Date are reported by test specimen collection date; prior to this date, these were reported by test result date.

We are working to backfill all the datasets with the entire history prior to the dates above and make them available here for use by the open data community.

Questions about these datasets? Visit the OpenDataPhilly Discussion Group: https://groups.google.com/forum/#!forum/opendataphilly

For terms of use: https://www.opendataphilly.org/organization/about/city-of-philadelphia

---------

### How to use
To get started reading and analyzing the data:

* Clone this repo
* If you use R/RStudio, open a new project based on an existing folder, pointing to the local copy of the repo
* Read all the files within a folder as a historical dataset, for example:
```{r message=FALSE, warning=FALSE}
library(tidyverse)

# Stack daily files within folder and keep distinct records
build_historical_dataset <- function(data_folder) {
  list.files(path = data_folder, full.names = TRUE) %>%
    map_dfr(read_csv) %>%
    distinct() # duplicates occur when there is an extract but no data update
}

# Cases by test result and reporting dates
cases_by_date <- build_historical_dataset("cases_by_date")
cases_by_date

# Cases by zip code and reporting date
cases_by_zipcode <- build_historical_dataset("cases_by_zipcode")
cases_by_zipcode
```


---------

### Analysis: COVID-19's incidence and effective reproductive number in Philly
```{r message=FALSE, warning=FALSE}
library(lubridate)
library(EpiEstim)

# Daily case count by test result date
incidence_data <- list.files(path = "cases_by_date", full.names = TRUE) %>% 
  last %>% 
  read_csv(col_types = cols(collection_date = col_character(),
                            etl_timestamp = col_skip(),
                            negative = col_integer(),
                            positive = col_integer())) %>% 
  filter(!is.na(positive) & (date(collection_date) >= date("2020-03-16"))) %>%
  mutate(dates = date(collection_date)) %>%
  arrange(dates) %>%
  mutate(positivity_rate = positive / (positive + negative)) %>%
  select(dates, positive, negative, positivity_rate) %>%
  filter(dates <= last(dates) - 3) # remove last 3 days considering lag in test results

# Plot incidence and effective reproductive number over time
# Serial interval mean and std estimates from: https://www.dhs.gov/publication/st-master-question-list-covid-19
res_parametric_si <- estimate_R(incidence_data %>% 
                                  select(dates, I = positive),
                                method="parametric_si",
                                config = make_config(list(mean_si = 5.29, std_si = 5.32)))
plot(res_parametric_si, legend = FALSE)
```
