---
title: "covid19-philadelphia"
output: github_document
---
De-identified, aggregate datasets showing COVID-19 cases and deaths by date, zip, or age/sex as made available by the City of Philadelphia through its Open Data Program.

The data in this repository (collected since 6/4/20) is a history of daily snapshots made available here: https://www.opendataphilly.org/showcase/philadelphia-covid-19-information

For terms of use: https://www.opendataphilly.org/organization/about/city-of-philadelphia

---------
If you use R/RStudio, to get started reading and analyzing the data:

* Clone this repo
* Open a new project with RStudio based on an existing folder, pointing to the local copy of the repo
* Read all the files within a folder as a historical dataset:
```{r message=FALSE, warning=FALSE}
library(tidyverse)

# Stack daily files within folder and keep distinct records
build_historical_dataset <- function(data_folder) {
  list.files(path = data_folder, full.names = TRUE) %>%
    map_dfr(read_csv) %>%
    distinct() # duplicates occur when there is an extract but no data update
}
cases_by_date <- build_historical_dataset("cases_by_date")
cases_by_date
```

* Explore the epidemic's incidence and effective reproductive number:
```{r message=FALSE, warning=FALSE}
library(lubridate)
library(EpiEstim)

# Daily case count by test result date
incidence_data <- list.files(path = "cases_by_date", full.names = TRUE) %>% 
  last %>% 
  read_csv(col_types = cols(result_date = col_character(),
                            etl_timestamp = col_skip(),
                            negative = col_integer(),
                            positive = col_integer())) %>% 
  filter(!is.na(positive) & (date(result_date) >= date("2020-03-16"))) %>%
  mutate(dates = date(result_date)) %>%
  arrange(dates) %>%
  mutate(positivity_rate = positive / (positive + negative)) %>%
  mutate(positivity_rate_last = lag(positivity_rate, default = first(positivity_rate))) %>%
  select(dates, positive, negative, positivity_rate, positivity_rate_last) %>%
  filter(dates <= last(dates) - 3) # remove last 3 days considering lag in test results

# Plot incidence and effective reproductive number over time
# Serial interval mean and std estimates from: https://www.dhs.gov/publication/st-master-question-list-covid-19
res_parametric_si <- estimate_R(incidence_data %>% 
                                  select(dates, I = positive),
                                method="parametric_si",
                                config = make_config(list(mean_si = 5.29, std_si = 5.32)))
plot(res_parametric_si, legend = FALSE)
```
